Diffusive Homeostasis in a Self-Organizing Recurrent Neural Network \\ \begin{large} Spatially dependent Interaction as a Determinant of Neural Activity and Plasticity \end{large}

### 2017-03-13 draft.tex
.

\clubpenalty = 1000 % schliesst Schusterjungen aus
\widowpenalty = 1000 \displaywidowpenalty = 1000% schliesst Hurenkinder aus

@@

\tableofcontents
\newpage
\clearpage

@@

Theoretical models of homeostasis are generally based on a form of feedback control. This is also known as a \textit{closed loop control} in control theory. All these systems include a number of necessary functional components, which will be briefly introduced in the following. Concepts and terms were taken from \cite[p. 59-64]{Cybernetik_Systems_Cruse_2006}.
\newpage
\subsubsection{Theory of Control Systems} \label{Control_Sys_Theo_Section}
A closed loop control generally must contain some \emph{process} that is to be regulated. Despite its potential internal complexity, its functional role in the context of a feedback system is to simply generate a time-dependent output based on the history of a time-dependent input. This output is what one wishes to control and regulate by means of the feedback loop and thus represents the signal that is further processed. It might not be possible however to directly quantify this desired output in terms of some internal state of the process. The firing rate of a neuron for example is an abstract quantity, which can not be \emph{directly} related to some neuronal state variable. Consequently, it might be necessary to transform the actual output of the process into a different signal by a \textit{feedback transducer}. The (potentially transformed) output is fed into a \textit{comparator} in the next step, which---unsurprisingly---compares the given output of the process to some given target or \textit{reference}, typically by subtraction. Note that this reference can but does note have to be a time-dependent signal. The term \textit{homeostasis} however generally refers to a fixed reference, while implementations with a time dependent reference can be regarded as a form of \textit{servomechanism}. The difference between reference and the feedback transducer's signal is known as the \textit{error} signal. This error signal is given to a \textit{controller}, which transforms this error signal into an appropriate input signal for the initially introduced process. This is also the point in the control loop where \textit{external disturbances} come into play by being added to the input generated by the controller. We illustrated the described flow of feedback control in Figure \ref{Closed_Loop_Control_Illustration}.
\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{../../plots/illustrations/closed_loop_control.eps}
\end{center}
\caption[Schematic illustration of a closed loop control]{Schematic illustration of a closed loop control as described in Section \ref{Control_Sys_Theo_Section}}
\label{Closed_Loop_Control_Illustration}
\end{figure}
Based on the way the controller generates input for the dynamic process based on the error signal one can classify control loops by three types:
\newpage
\begin{itemize}
\item A \textit{proportional controller} generates a signal that is proportional to the error signal it receives. The most important property of a proportional controller is that the stationary state of the control loop generally does not imply an error signal equal to zero.
\item An \textit{integrator controller} uses an integration over the past of the error signal to obtain its output. The stationary solution necessarily implies that the error signal is zero.
\item A \textit{differential controller} generates its output by means of the time derivative of the error signal. This type of controller can quickly react to momentary changes but does not compensate for sustained deviations of the error signal from zero.
\end{itemize}
It should be pointed out that these types of control mechanisms can be simultaneously used in a control system, e.g. through a linear combination. Furthermore, the presented terms refer to concepts of linear control theory. In the study of nonlinear feedback systems they thus only apply in a limited regime where a first order approximation of the system is sufficiently exact.

@@

\begin{table}
\caption{Parameters of synaptic connections.}
\begin{tabu}{|l|l|l|l|l|}
\hline
\textbf{parameter} & \textbf{EE} & \textbf{EI} & \textbf{IE} & \textbf{II} \\ \hline
connection fraction & $\rightarrow 0.1$ & $0.1$ & $0.1$ & $0.5$ \\ \hline
initial connection strength & $0.0001 mV$ & $1.5 mV$ & $-1.5 mV$ & $-1.5 mV$ \\ \hline
conduction delay & $1.5 ms$ & $0.5 ms$ & $1.0 ms$ & $1.0 ms$ \\
\hline
\end{tabu}
\label{syn_conn_params}
\end{table}
\newpage

@@

\newpage

@@

\newpage
\section{Results} \label{results}
We first present the results of a comparison between features of the network activity in both homeostatic variants in Section \ref{activ_analys}, since our main goal of the implementation of diffusive homeostasis was to allow the network to develop a broader distribution of firing rates across excitatory neurons compared to the original version of homeostasis. This is followed by an analytic discussion of an instability we observed within the diffusive homeostatic feedback loop, see Section \ref{theor_osc}. Furthermore, we compare topological features of the network under the influence of diffusive and non-diffusive homeostasis in section \ref{Network_Topology_Section}.

@@

Returning to the question of how to describe NO synthesis in a simplified form, we note that according to Equation \eqref{nNOS_single_sol_approx2}, a single spike causes the release of NO by an amount of $\frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}}\mathrm{ln}(2)}{3}$, which makes the mean rate of NO production over time simply $\langle nNOS \rangle = \langle r \rangle \frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}} \mathrm{ln} (2)}{3}$. However, it is not sufficient to simply propose $nNOS_i(t) = r_i(t) \frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}} \mathrm{ln} (2)}{3}$. A single cell fires in the range of $\mathrm{3\,Hz}$. Therefore, the dynamics of $\rm nNOS$ are dominated by single spike events as shown in Figure \ref{nNOS_approx_plot} and it is---at least on the time scale of the observed oscillations---not a proper approximation to describe $nNOS$ in this smooth, rate-based fashion.
\newpage
We introduced an idealized picture of diffusive homeostasis to deal with this, where diffusion across the tissue is instantaneous. This simplification results in a single level of NO-concentration for all neurons, only being modified over time by the sum of all neurons' NO-syntheses $nNOS_{\rm total}(t) \equiv \sum_{i} nNOS_i(t)$ and the decay term $-\lambda NO$. $nNOS_{\rm total}$ must be divided by the area $L^2$ of the tissue to account for the spread of NO. Furthermore, we introduce a random fluctuation term $\sigma_{NO} \xi(t)$ that accounts for local, momentary deviations from the mean. The idealized version of \eqref{NO_dyn} then reads
\begin{equation}
\dot{NO}(t)=-\lambda NO + \frac{nNOS_{\rm total}(t)}{L^2} + \sigma_{NO} \xi(t)\;.
\label{NO_dyn_approx}
\end{equation}
The amount of noise in the system is a hard-to-predict quantity since it depends on the spatial structure on the diffusive lattice, giving neurons that are further separated less impact onto each other. We tried to estimate the noise level based on the same instant-spread simplification made in Equation \eqref{NO_dyn_approx} to describe the overall NO-dynamics. This led to an underestimation of the total energy within the power spectrum in the final result. However, the noise level only acts as a global scaling factor of the shape of the power spectrum as long as we assume the noise in the system to be approximately white and was thus left open as a model fitting parameter.

@@

\subsubsection{Excitatory Connection Fraction}
In the previous version of the LIF-SORN, initializing the network with zero recurrent excitatory connections led to a overall increasing but saturating time course of the excitatory-to-excitatory connection fraction (CF). The growth rate was tuned such that the terminal CF settled at $\mathrm{10\%}$, see \cite[p. 8]{SORN_Paper}. Omitting the distance dependence of connection probability for synaptic growth led to a slightly higher connection fraction. We ran a full simulation under diffusive homeostasis to compare these previous results to the diffusive case, including the growth phase by setting the target concentration to an appropriate value that was determined in a previous simulation with non-diffusive homeostasis. The resulting time course of CF is depicted in Figure \ref{CF_plot}.
\begin{figure}
%%% plot_CF.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/CF_plot.png}
\caption[Connection fractions of recurrent excitatory synapses]{Connection fractions of recurrent excitatory synapses. Simulation protocols include all four possible combinations of diffusive/non-diffusive homeostasis and distance-dependent/non-distance-dependent synaptic growth, each variant retained for the entire simulation time. Data from 10 trials, linewidth represents standard error.}
\label{CF_plot}
\end{figure}
Diffusive homeostasis caused the CF to slightly overshoot its final fraction in the growth phase. To understand this kind of behavior, we recalled the two mechanisms that directly determine changes within the connection fraction, namely synaptic creation and pruning. The average change of connection fraction is simply proportional to the difference between newly created synapses and those that were removed. Though fluctuating, the synaptic growth rate was kept at a constant level of 920 synapses/second. Any differences in Figure \ref{CF_plot} must therefore originate in different pruning rates. Figure \ref{STDP_change} shows the mean and standard deviation of weight changes due to STDP between two normalization steps. Non-diffusive homeostasis has a higher mean as well as a broader distribution of weight fluctuations throughout the beginning of the simulation. However, the difference of mean changes is orders of magnitude smaller than the width of fluctuations. Thus, while it might seem counter-intuitive that a higher positive mean weight change comes with a smaller chance of synaptic survival, we attribute the main cause of overshooting in Figure \ref{CF_plot} to the increased fluctuation of weights since it raises the chance of going below the pruning threshold within a given time interval.
\begin{figure}
%%% stdp_change_time.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/av_stdp_change_time.png}
\caption[Mean and standard deviation of weight changes between synaptic normalization]{Top: Mean change of synaptic weights (creation and pruning excluded) between two normalization steps. Bottom: Standard deviation for the same data. Both curves correspond to those in Figure \ref{CF_plot} with the same color.}
\label{STDP_change}
\end{figure}

@@

In the plastic network, the absence of this breaking of spatial homogeneity even led to an under-representation of bidirectional connections, which is known to be an effect of STDP in recurrent networks \cite{Syn_Plast_Abbott}. The separation between simulations with and without a spatial connection profile is retained after having implemented diffusive homeostasis, as shown in Figure \ref{RC_CFB_plot}. 
\begin{figure}
%%% plot_RC_CFB.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/RC_CFB_plot.png}
\caption[Fraction of bidirectional connections]{Fraction of bidirectional connections normalized over the expected fraction for a random graph with equal total connection fraction. Data from 10 trials, linewidth represents standard error.}
\label{RC_CFB_plot}
\end{figure}
\subsubsection{Distribution of Synaptic Weights} \label{Syn_Weight_Dist_Section}
A well-studied property of neural networks is the distribution of synaptic weights. Experimental studies in hippocampus and cortical regions suggest heavy-tailed log-normal-like distributions of excitatory synaptic efficacies \cite{Song_Connectivity_2005,Lisman_Synapses_1993,Yasumatsu_Synapses_2008,Loewenstein_Spine_Sizes}. Theoretical models are mainly based on a combination of multiplicative and additive weight dynamics \cite{Loewenstein_Spine_Sizes,Statman_Synapses_2014}, which is in line with our implementation of multiplicative normalization and additive STDP. Figure \ref{Weight_Dist} compares the resulting distributions. We found that we can retain a log-normal-like distribution in combination with diffusive homeostasis, giving a similar distribution as for non-diffusive homeostasis. Both distributions were slightly left-skewed ($\mathrm{v_{\rm diff.} = -1.21}$,$\mathrm{v_{\rm non-diff.} = -1.44}$). 
\begin{figure}
%%% plot_weight_dist_mult_sets.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/syn_weight_dist_mult_sets.png}
\caption[Distribution of decadic logarithm of excitatory synaptic weights]{Distribution of decadic logarithm of excitatory synaptic weights at $\mathrm{t=1500\,s}$ for diffusive and non-diffusive homeostasis (both with distance-dependent connectivity). Gaussian fits (non-linear least squares) returned $\mathrm{\mu_{\rm diff.}=-0.128\, \mathrm{log}_{10}(mV)}$, $\mathrm{\mu_{\rm non-diff.}=-0.064\, \mathrm{log}_{10}(mV)}$, $\mathrm{\sigma_{\rm diff.}=0.415\, log_{10}(mV)}$ and $\mathrm{\sigma_{non-diff.}=0.395\, log_{10}(mV)}$. Weight matrices were taken from 5 simulations, each at $\rm t = 1500\,s$.}
\label{Weight_Dist}
\end{figure}

@@

\begin{table}
\caption[p-values of the Kolmogorov-Smirnov test calculated for a subset of dataset combinations depicted in Figure \ref{Out_Weight_Mean_Quantile}]{p-values of the Kolmogorov-Smirnov test calculated for a subset of dataset combinations depicted in Figure \ref{Out_Weight_Mean_Quantile}. The abbreviations are as follows: D: diffusive, T: topology, S: shuffled weights, N: negotiation.}
\begin{tabu}{|l|l|}
\hline
\textbf{Combination} & \textbf{p-value} \\ \hline
D,T,NS vs ND,T,NS & $\rm 4.662 \cdot 10^{-35}$ \\ \hline
D,NT,NS vs ND,NT,NS & $\rm 6.895 \cdot 10^{-77}$ \\ \hline
D,T,NS vs D,T,S & $\rm 3.344 \cdot 10^{-44}$ \\ \hline
D,NT,NS vs D,NT,S & $\rm 1.790 \cdot 10^{-61}$ \\ \hline
\end{tabu}
\label{Kolmogorov_Mean_Out}
\end{table}

@@

\newpage
\section{Further Analysis of the Firing Rate Distribution}
The results of section \ref{Fir_Dist_Section} were positive on a preliminary level regarding our goal of achieving a broad and skewed firing rate distribution. Following this, we tried to further gain insight into the actual shape of the resulting distribution. 
\subsection{Review of a Dynamic Mean-Field Model}\label{Mean_Field_Review_Section}
Sweeney et al. attempted to predict the shape of the firing rate distribution by means of a simplified mean-field approximation of diffusion \cite{Sweeney_Paper}. The main simplification within homeostatic feedback was the replacement of Equation \ref{Theta_dyn} by
\begin{equation}
\dot{V_{t,i}}(t) = \frac{1}{\tau_{HIP}} \left( (1-\alpha)\frac{\phi_i-\phi_0}{\phi_i} +\alpha \frac{\langle \phi \rangle-\phi_0}{\langle \phi \rangle} \right)\;. 
\label{diff_hom_simpl_sweeney}
\end{equation}
$\mathrm{\alpha \: \epsilon \: [0,1]}$ thereby acted as a parameter that determines the mixture between single-neuron homeostasis ($\mathrm{\alpha=0}$) and the limit of quasi instantaneous spreading of the diffusive signal across the population ($\mathrm{\alpha=1}$). $\mathrm{\phi_i}$ and $\mathrm{\mathrm{\langle \phi \rangle}}$ represent individual firing rates and their population mean.

@@

\newpage
\subsection{Spreading of NO-sources extends Encoding of Input Heterogeneity}\label{NO_source_spread_section}
We considered the possibility that a wider spatial spread of each individual NO source---in contrast to the approximate point source used up to this point---could locally decrease distinguishability between neurons since our analytic results revealed that a strong correlation between spatial structure and resulting firing rates is present if diffusive interaction allows a clear distinction between individual neurons and their spatial neighbors. Referring to the results shown in Figure \ref{Ext_Input_Switch}, this modification with respect to the NO influx might allow input heterogeneities to be encoded over a longer period of time without the need to tweak diffusion to unrealistically high values. More specifically, one could potentially observe the distinction of input groups shown in Figure \ref{Ext_Input_Switch} over a longer time scale.  While Sweeney et al. argued that the size of the studied tissue is large compared to the size of individual somata (which, apparently, they claim to be the main source of nitric oxide) Philippides et al. reported that NO is often synthesized in a more delocalized way by means of fine fibers exceeding the soma \cite{Philippides_2005}. The diameter of these ``production areas" is on the order of $\mathrm{10-100 \mu m}$.

@@

\newpage
\section{Conclusion and Outlook}\label{Conclusion_Section}
Our main goal of implementing diffusive homeostasis was to allow a broad distribution of firing rates across the excitatory population of the LIF-SORN, since experimental research has shown this to be a extensively observable phenomenon in the cortex. The implementation of diffusive homeostasis can be regarded as a success with respect to this feature. Furthermore, we did not observe any breakdown of previously studied desirable features of network topology, emerging from the plasticity mechanisms included.

>>>>

\clubpenalty = 10000 % schliesst Schusterjungen aus
\widowpenalty = 10000 \displaywidowpenalty = 10000% schliesst Hurenkinder aus

@@

\tableofcontents

@@

Theoretical models of homeostasis are generally based on a form of feedback control. This is also known as a \textit{closed loop control} in control theory. All these systems include a number of necessary functional components, which will be briefly introduced in the following. Concepts and terms were taken from \cite[p. 59-64]{Cybernetik_Systems_Cruse_2006}.

\subsubsection{Theory of Control Systems} \label{Control_Sys_Theo_Section}
A closed loop control generally must contain some \emph{process} that is to be regulated. Despite its potential internal complexity, its functional role in the context of a feedback system is to simply generate a time-dependent output based on the history of a time-dependent input. This output is what one wishes to control and regulate by means of the feedback loop and thus represents the signal that is further processed. It might not be possible however to directly quantify this desired output in terms of some internal state of the process. The firing rate of a neuron for example is an abstract quantity, which can not be \emph{directly} related to some neuronal state variable. Consequently, it might be necessary to transform the actual output of the process into a different signal by a \textit{feedback transducer}. The (potentially transformed) output is fed into a \textit{comparator} in the next step, which---unsurprisingly---compares the given output of the process to some given target or \textit{reference}, typically by subtraction. Note that this reference can but does note have to be a time-dependent signal. The term \textit{homeostasis} however generally refers to a fixed reference, while implementations with a time dependent reference can be regarded as a form of \textit{servomechanism}. The difference between reference and the feedback transducer's signal is known as the \textit{error} signal. This error signal is given to a \textit{controller}, which transforms this error signal into an appropriate input signal for the initially introduced process. This is also the point in the control loop where \textit{external disturbances} come into play by being added to the input generated by the controller. We illustrated the described flow of feedback control in Figure \ref{Closed_Loop_Control_Illustration}.
\begin{figure}
\begin{center}
\includegraphics[width=0.7\textwidth]{../../plots/illustrations/closed_loop_control.eps}
\end{center}
\caption[Schematic illustration of a closed loop control]{Schematic illustration of a closed loop control as described in Section \ref{Control_Sys_Theo_Section}}
\label{Closed_Loop_Control_Illustration}
\end{figure}
Based on the way the controller generates input for the dynamic process based on the error signal one can classify control loops by three types:
\begin{itemize}
\item A \textit{proportional controller} generates a signal that is proportional to the error signal it receives. The most important property of a proportional controller is that the stationary state of the control loop generally does not imply an error signal equal to zero.
\item An \textit{integrator controller} uses an integration over the past of the error signal to obtain its output. The stationary solution necessarily implies that the error signal is zero.
\item A \textit{differential controller} generates its output by means of the time derivative of the error signal. This type of controller can quickly react to momentary changes but does not compensate for sustained deviations of the error signal from zero.
\end{itemize}
It should be pointed out that these types of control mechanisms can be simultaneously used in a control system, e.g. through a linear combination. Furthermore, the presented terms refer to concepts of linear control theory. In the study of nonlinear feedback systems they thus only apply in a limited regime where a first order approximation of the system is sufficiently exact.

@@

\begin{table}
\caption{Parameters of synaptic connections.}
\begin{tabu}{|l|l|l|l|l|}
\hline
\textbf{parameter} & \textbf{EE} & \textbf{EI} & \textbf{IE} & \textbf{II} \\ \hline
connection fraction & $\rightarrow 0.1$ & $0.1$ & $0.1$ & $0.5$ \\ \hline
initial connection strength & $0.0001 mV$ & $1.5 mV$ & $-1.5 mV$ & $-1.5 mV$ \\ \hline
conduction delay & $1.5 ms$ & $0.5 ms$ & $1.0 ms$ & $1.0 ms$ \\
\hline
\end{tabu}
\label{syn_conn_params}
\end{table}

@@

@@

\section{Results} \label{results}
We first present the results of a comparison between features of the network activity in both homeostatic variants in Section \ref{activ_analys}, since our main goal of the implementation of diffusive homeostasis was to allow the network to develop a broader distribution of firing rates across excitatory neurons compared to the original version of homeostasis. This is followed by an analytic discussion of an instability we observed within the diffusive homeostatic feedback loop, see Section \ref{theor_osc}. Furthermore, we compare topological features of the network under the influence of diffusive and non-diffusive homeostasis in section \ref{Network_Topology_Section}.

@@

Returning to the question of how to describe NO synthesis in a simplified form, we note that according to Equation \eqref{nNOS_single_sol_approx2}, a single spike causes the release of NO by an amount of $\frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}}\mathrm{ln}(2)}{3}$, which makes the mean rate of NO production over time simply $\langle nNOS \rangle = \langle r \rangle \frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}} \mathrm{ln} (2)}{3}$. However, it is not sufficient to simply propose $nNOS_i(t) = r_i(t) \frac{{Ca^{2+}_{\rm spike}}^3 \tau_{Ca^{2+}} \mathrm{ln} (2)}{3}$. A single cell fires in the range of $\mathrm{3\,Hz}$. Therefore, the dynamics of $\rm nNOS$ are dominated by single spike events as shown in Figure \ref{nNOS_approx_plot} and it is---at least on the time scale of the observed oscillations---not a proper approximation to describe $nNOS$ in this smooth, rate-based fashion.

We introduced an idealized picture of diffusive homeostasis to deal with this, where diffusion across the tissue is instantaneous. This simplification results in a single level of NO-concentration for all neurons, only being modified over time by the sum of all neurons' NO-syntheses $nNOS_{\rm total}(t) \equiv \sum_{i} nNOS_i(t)$ and the decay term $-\lambda NO$. $nNOS_{\rm total}$ must be divided by the area $L^2$ of the tissue to account for the spread of NO. Furthermore, we introduce a random fluctuation term $\sigma_{NO} \xi(t)$ that accounts for local, momentary deviations from the mean. The idealized version of \eqref{NO_dyn} then reads
\begin{equation}
\dot{NO}(t)=-\lambda NO + \frac{nNOS_{\rm total}(t)}{L^2} + \sigma_{NO} \xi(t)\;.
\label{NO_dyn_approx}
\end{equation}
The amount of noise in the system is a hard-to-predict quantity since it depends on the spatial structure on the diffusive lattice, giving neurons that are further separated less impact onto each other. We tried to estimate the noise level based on the same instant-spread simplification made in Equation \eqref{NO_dyn_approx} to describe the overall NO-dynamics. This led to an underestimation of the total energy within the power spectrum in the final result. However, the noise level only acts as a global scaling factor of the shape of the power spectrum as long as we assume the noise in the system to be approximately white and was thus left open as a model fitting parameter.

@@

\subsubsection{Excitatory Connection Fraction}
In the previous version of the LIF-SORN, initializing the network with zero recurrent excitatory connections led to a overall increasing but saturating time course of the excitatory-to-excitatory connection fraction (CF). The growth rate was tuned such that the terminal CF settled at $\mathrm{10\%}$, see \cite[p. 8]{SORN_Paper}. Omitting the distance dependence of connection probability for synaptic growth led to a slightly higher connection fraction. We ran a full simulation under diffusive homeostasis to compare these previous results to the diffusive case, including the growth phase by setting the target concentration to an appropriate value that was determined in a previous simulation with non-diffusive homeostasis. The resulting time course of CFs is depicted in Figure \ref{CF_plot}.
\begin{figure}
%%% plot_CF.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/CF_plot.png}
\caption[Connection fractions of recurrent excitatory synapses]{Connection fractions of recurrent excitatory synapses. Simulation protocols include all four possible combinations of diffusive/non-diffusive homeostasis and distance-dependent/non-distance-dependent synaptic growth, each variant retained for the entire simulation time. Data from 5 trials, linewidth represents standard error.}
\label{CF_plot}
\end{figure}
It appeared to be the case that diffusive homeostasis caused the CF to slightly overshoot its non-diffusive counterpart in the growth phase. To estimate statistical significance, we performed a \textit{two-sample t-test} for the sample data taken from $\rm t = 300\, s$, which---judging by eye---is the point where the overshoot of CFs resulting from diffusive homeostasis is at its maximum. The null hypothesis of the two-sample t-test states that the probability density functions underlying two sets of sample data have the same expected value, under the assumption that both are normally distributed \cite{EncyMath_t_test} and have equal variance. Since we could not guarantee the last assumption, we used a generalization by Welch that allows for different variances \cite{Welch_1947}. The condition of normality was left as an assumption. Comparing diffusive versus non-diffusive data, we found a p-value of $p_{\rm topol.} = \rm 1.25 \cdot 10^{-3}$ for networks with a spatially-dependent connection profile and $p_{\rm no topol.} = \rm 4.00 \cdot 10^{-3}$ for networks with a flat connection profile. In terms of usual significance levels (1--10\% \cite{EncyMath_sign_level}), we thus rejected the null hypothesis and concluded that the observed overshoot is statistically significant for both network-growth variants.

To understand this overshoot, we recalled the two mechanisms that directly determine changes within the connection fraction, namely synaptic creation and pruning. The average change of connection fraction is simply proportional to the difference between newly created synapses and those that were removed. Though fluctuating, the synaptic growth rate was kept at a constant level of 920 synapses/second. Any differences in Figure \ref{CF_plot} must therefore originate in different pruning rates. Figure \ref{STDP_change} shows the mean and standard deviation of weight changes due to STDP between two normalization steps. Non-diffusive homeostasis has a higher mean as well as a broader distribution of weight fluctuations throughout the beginning of the simulation. However, the difference of mean changes is orders of magnitude smaller than the width of fluctuations. Thus, while it might seem counter-intuitive that a higher positive mean weight change comes with a smaller chance of synaptic survival, we attribute the main cause of overshooting in Figure \ref{CF_plot} to the increased fluctuation of weights since it raises the chance of going below the pruning threshold within a given time interval.
\begin{figure}
%%% stdp_change_time.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/av_stdp_change_time.png}
\caption[Mean and standard deviation of weight changes between synaptic normalization]{Top: Mean change of synaptic weights (creation and pruning excluded) between two normalization steps. Bottom: Standard deviation for the same data. Both curves correspond to those in Figure \ref{CF_plot} with the same color.}
\label{STDP_change}
\end{figure}

@@

In the plastic network, the absence of this breaking of spatial homogeneity even led to an under-representation of bidirectional connections, which is known to be an effect of STDP in recurrent networks \cite{Syn_Plast_Abbott}. The separation between simulations with and without a spatial connection profile is retained after having implemented diffusive homeostasis, as shown in Figure \ref{RC_CFB_plot}. 
\begin{figure}
%%% plot_RC_CFB.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/RC_CFB_plot.png}
\caption[Fraction of bidirectional connections]{Fraction of bidirectional connections normalized over the expected fraction for a random graph with equal total connection fraction. Data from 5 trials, linewidth represents standard error.}
\label{RC_CFB_plot}
\end{figure}
\subsubsection{Distribution of Synaptic Weights} \label{Syn_Weight_Dist_Section}
A well-studied property of neural networks is the distribution of synaptic weights. Experimental studies in hippocampus and cortical regions suggest heavy-tailed log-normal-like distributions of excitatory synaptic efficacies \cite{Song_Connectivity_2005,Lisman_Synapses_1993,Yasumatsu_Synapses_2008,Loewenstein_Spine_Sizes}. Theoretical models are mainly based on a combination of multiplicative and additive weight dynamics \cite{Loewenstein_Spine_Sizes,Statman_Synapses_2014}, which is in line with our implementation of multiplicative normalization and additive STDP. Figure \ref{Weight_Dist} compares the resulting distributions. We found that we can retain a log-normal-like distribution in combination with diffusive homeostasis, giving a similar distribution as for non-diffusive homeostasis. Both distributions were slightly left-skewed ($\mathrm{v_{\rm diff.} = -1.21}$,$\mathrm{v_{\rm non-diff.} = -1.44}$). 
\begin{figure}
%%% plot_weight_dist_mult_sets.py
\includegraphics[width=\textwidth]{../../plots/syn_topology/syn_weight_dist_mult_sets.png}
\caption[Distribution of decadic logarithm of excitatory synaptic weights]{Distribution of decadic logarithm of excitatory synaptic weights at $\mathrm{t=1500\,s}$ for diffusive and non-diffusive homeostasis (both with distance-dependent connectivity). Gaussian fits (non-linear least squares) returned $\mathrm{\mu_{\rm diff.}=-0.128\, \mathrm{log}_{10}(mV)}$, $\mathrm{\mu_{\rm non-diff.}=-0.064\, \mathrm{log}_{10}(mV)}$, $\mathrm{\sigma_{\rm diff.}=0.415\, log_{10}(mV)}$ and $\mathrm{\sigma_{non-diff.}=0.395\, log_{10}(mV)}$. Weight matrices were taken from 5 simulations, each at $\rm t = 1500\,s$.}
\label{Weight_Dist}
\end{figure}

@@

\begin{table}
\caption[p-values of the Kolmogorov-Smirnov test calculated for a subset of dataset combinations depicted in Figure \ref{Out_Weight_Mean_Quantile}]{p-values of the Kolmogorov-Smirnov test calculated for a subset of dataset combinations depicted in Figure \ref{Out_Weight_Mean_Quantile}. Abbreviations are as follows: D: diffusive, T: topology, S: shuffled weights, N: negotiation.}
\begin{tabu}{|l|l|}
\hline
\textbf{Combination} & \textbf{p-value} \\ \hline
D,T,NS vs ND,T,NS & $\rm 4.662 \cdot 10^{-35}$ \\ \hline
D,NT,NS vs ND,NT,NS & $\rm 6.895 \cdot 10^{-77}$ \\ \hline
D,T,NS vs D,T,S & $\rm 3.344 \cdot 10^{-44}$ \\ \hline
D,NT,NS vs D,NT,S & $\rm 1.790 \cdot 10^{-61}$ \\ \hline
\end{tabu}
\label{Kolmogorov_Mean_Out}
\end{table}

@@

\section{Further Analysis of the Firing Rate Distribution}
The results of section \ref{Fir_Dist_Section} were positive on a preliminary level regarding our goal of achieving a broad and skewed firing rate distribution. Following this, we tried to further gain insight into the actual shape of the resulting distribution. 
\subsection{Review of a Dynamic Mean-Field Model}\label{Mean_Field_Review_Section}
Sweeney et al. attempted to predict the shape of the firing rate distribution by means of a simplified mean-field approximation of diffusion \cite{Sweeney_Paper}. The main simplification within homeostatic feedback was the replacement of Equation \ref{Theta_dyn} by
\begin{equation}
\dot{V_{t,i}}(t) = \frac{1}{\tau_{HIP}} \left( (1-\alpha)\frac{\phi_i-\phi_0}{\phi_i} +\alpha \frac{\langle \phi \rangle-\phi_0}{\langle \phi \rangle} \right)\;. 
\label{diff_hom_simpl_sweeney}
\end{equation}
$\mathrm{\alpha \: \epsilon \: [0,1]}$ thereby acted as a parameter that determines the mixture between single-neuron homeostasis ($\mathrm{\alpha=0}$) and the limit of quasi instantaneous spreading of the diffusive signal across the population ($\mathrm{\alpha=1}$). $\mathrm{\phi_i}$ and $\mathrm{\mathrm{\langle \phi \rangle}}$ represent individual firing rates and their population mean.

@@

\subsection{Spreading of NO-sources extends Encoding of Input Heterogeneity}\label{NO_source_spread_section}
We considered the possibility that a wider spatial spread of each individual NO source---in contrast to the approximate point source used up to this point---could locally decrease distinguishability between neurons since our analytic results revealed that a strong correlation between spatial structure and resulting firing rates is present if diffusive interaction allows a clear distinction between individual neurons and their spatial neighbors. Referring to the results shown in Figure \ref{Ext_Input_Switch}, this modification with respect to the NO influx might allow input heterogeneities to be encoded over a longer period of time without the need to tweak diffusion to unrealistically high values. More specifically, one could potentially observe the distinction of input groups shown in Figure \ref{Ext_Input_Switch} over a longer time scale.  While Sweeney et al. argued that the size of the studied tissue is large compared to the size of individual somata (which, apparently, they claim to be the main source of nitric oxide) Philippides et al. reported that NO is often synthesized in a more delocalized way by means of fine fibers exceeding the soma \cite{Philippides_2005}. The diameter of these ``production areas" is on the order of $\mathrm{10-100 \mu m}$.

@@

\section{Conclusion and Outlook}\label{Conclusion_Section}
Our main goal of implementing diffusive homeostasis was to allow a broad distribution of firing rates across the excitatory population of the LIF-SORN, since experimental research has shown this to be a extensively observable phenomenon in the cortex. The implementation of diffusive homeostasis can be regarded as a success with respect to this feature. Furthermore, we did not observe any breakdown of previously studied desirable features of network topology, emerging from the plasticity mechanisms included.
